A content-sensitive framework for modeling episodic memory reveals event-like structure in naturalistic experience, recall, and neural processing

### 2019-07-24 paper/main.tex
finished revision

%\title{How is experience transformed into memory?}
%\title{A content-based framework for capturing stimulus and memory dynamics reveals event-like structure in naturalistic episodic recall}
%\title{A content-based model characterizing/describing/representing how experiences and memories unfold over time reveals event-like structure in naturalistic episodic recall}
%\title{A content-based framework describing the temporal dynamics of experiences and memories reveals event-like structure in naturalistic episodic recall}
%\title{A content-based model of how experiences and memories unfold over time reveals event-like structure in naturalistic episodic recall}
\title{A novel framework for linking dynamic experiences to memories reveals event-like structure in naturalistic episodic recall}
%\title{Recalling naturalistic experiences exhibits event-like structure}

@@

\begin{abstract}
Our life experiences unfold over time in highly complex manner, with the evolving presence and absence of numerous intricate features describing our journey between each circumstance or event we encounter. Here, we propose a framework for mapping dynamic naturalistic experiences onto geometric spaces as \textit{trajectories} that capture the temporal dynamics of real-world content. Within this geometric framework, one may compare the shape of the trajectory formed by an experience to that defined by one's later recollection to characterize our memories' recovery and distortion of the external world. Here, we apply this approach to a naturalistic memory experiment in which participants viewed and verbally recounted a video, and find that the video and subsequent recalls share both an experience-specific shape and a discernible event-like structure.  However, the level of \textit{precision} with which individuals recounted various events and the \textit{distinctiveness} of recall for those events were varied and predictive of overall memory performance. Finally, we identify a network of brain structures that is sensitive to the ``shapes" of our ongoing experiences, and an overlapping network sensitive to how we will later remember them.  These results highlight the rich event-like structure of the external world and our memories, and offer novel, content-sensitive alternatives to classical ``proportion recalled" measures for assessing episodic memory
%Our life experiences unfold over time and the temporal dynamics of their contents form unique \textit{experience trajectories}.  Within this geometric framework, one can compare the shape of the trajectory formed by an experience to that defined by our later remembering of that experience.  We propose a framework for mapping naturalistic experiences onto geometric spaces that characterize how they unfold over time.  We apply this approach to a naturalistic memory experiment which had participants view and recount a video.  We find that the video and subsequent recall share an event-like structure, and that the ``shapes'' of the trajectories formed by participants' recountings were all highly similar to that of the original video. Further, participants' that recounted the events with a high level of precision and in a distinctive way had better subsequent memory performance. Lastly, we identified a network of brain structures that are sensitive to the shapes of our ongoing experiences, and an overlapping network that is sensitive to how we will later remember those experiences. These results highlight the rich event-like structure of memories for our life experiences, and introduce a novel framework for mapping between dynamic life experiences and their mnemonic counterparts.
\end{abstract}

@@

What does it mean to \textit{remember} something? In traditional episodic memory experiments \citep[e.g., list-learning or trial-based experiments;][]{Murd62a, Kaha96}, remembering is often cast as a discrete and binary operation: each studied item may be separated from the rest of one's experiences, and that item may be labeled as having been recalled versus forgotten. More nuanced studies might incorporate self-reported confidence measures as a proxy for memory strength, or ask participants to discriminate between ``recollecting'' the (contextual) details of an experience or having a general feeling of ``familiarity'' \citep{Yone02}. Using well-controlled, trial-based experimental designs, the field has amassed a wealth of valuable information regarding human episodic memory.  However, there are fundamental properties of the external world and our memories that trial-based experiments are not well suited to capture~\citep[for review also see][]{KoriGold94, HukEtal18}.  First, our experiences and memories are continuous, rather than discrete---removing a (naturalistic) event from the context in which it occurs can substantially change its meaning.  Second, the specific language used to describe an experience has little bearing on whether the experience should be considered to have been ``remembered.''  Asking whether the rememberer has precisely reproduced a specific set of words to describe a given experience is nearly orthogonal to whether they were actually able to remember it.  In classic (e.g., list-learning) memory studies, by contrast, the number or proportion of precise recalls is often a primary metric for assessing the quality of participants' memories.  Third, one might remember the \textit{essence} (or a general summary) of an experience but forget (or neglect to recount) particular details.  Capturing the essence of what happened is typically the main ``point'' of recounting a memory to a listener, while the addition of highly specific details may add comparatively little to successful conveyance of an experience.

How might one go about formally characterizing the essence of an experience, or whether it has been recovered by the rememberer?  Any given moment of an experience derives meaning from surrounding moments, as well as from longer-range temporal associations~\citep[e.g., ][]{LernEtal11, Mann19}.  Therefore, the timecourse describing how an event unfolds is fundamental to its overall meaning.  Further, this hierarchy formed by our subjective experiences at different timescales defines a \textit{context} for each new moment~\citep[e.g., ][]{HowaKaha02a, HowaEtal14}, and plays an important role in how we interpret that moment and remember it later~\citep[for review see][]{MannEtal15}.  Our memory systems can leverage these associations to form predictions that help guide our behaviors~\citep{RangRitc12}.  For example, as we navigate the world, the features of our subjective experiences tend to change gradually (e.g., the room or situation we are in at any given moment is strongly temporally autocorrelated), allowing us to form stable estimates of our current situation and behave accordingly~\citep{ZackEtal07, ZwaaRadv98}.

@@

Here we sought to examine how the temporal dynamics of a ``naturalistic'' experience were later reflected in participants' memories.  We analyzed an open dataset that comprised behavioral and functional Magnetic Resonance Imaging (fMRI) data collected as participants viewed and then verbally recounted an episode of the BBC television series \textit{Sherlock}~\citep{ChenEtal17}.  We developed a computational framework for characterizing the temporal dynamics of the moment-by-moment content of the episode and of participants' verbal recalls.  Specifically, we use topic modeling~\citep{BleiEtal03} to characterize the thematic conceptual (semantic) content present in each moment of the episode and recalls, and Hidden Markov Models~\citep{Rabi89, BaldEtal17} to discretize the evolving semantic content into events.  In this way, we cast naturalistic experiences (and recalls of those experiences) as \textit{trajectories} that describe how the experiences evolve over time. Under this framework, successful remembering entails verbally ``traversing'' the content trajectory of the episode, thereby reproducing the shape (or essence) of the original experience.  Comparing the shapes of the topic trajectories of the episode and of participants' retellings of the episode then reveals which aspects of the episode were preserved (or lost) in the translation into memory.  We further examine whether 1) the \textit{precision} with which a participant recounts each event and 2) the \textit{distinctiveness} each recall event is (relative to the other recalled events) relates to their overall memory performance.  Last, we identify networks of  brain structures whose responses (as participants watched the episode) reflected the shape of the episode, and how participants would later recount the episode.

\section*{Results}
To characterize the shape of the \textit{Sherlock} episode and participants' subsequent recountings of its unfolding, we used a topic model~\citep{BleiEtal03} to discover the latent themes in the episode's dynamic content.  Topic models take as inputs a vocabulary of words to consider and a collection of text documents; they return as output two matrices.  The first output is a \textit{topics matrix} whose rows are topics (latent themes) and whose columns correspond to words in the vocabulary. The entries of the topics matrix define how each word in the vocabulary is weighted by each discovered topic.  For example, a detective-themed topic might weight heavily on words like ``crime,'' and ``search.''  The second output is a \textit{topic proportions matrix}, with one row per document and one column per topic.  The topic proportions matrix describes which mixture of topics is reflected in each document.

\cite{ChenEtal17} collected hand-annotated information about each of 1000 (manually identified) scenes spanning the roughly 50 minute video used in their experiment.  This information included: a brief narrative description of what was happening; whether the scene took place indoors vs. outdoors; names of any characters on the screen; names of any characters who were in focus in the camera shot; names of characters who were speaking; the location where the scene took place; the camera angle (close up, medium, long, etc.); whether or not background music was present; and other similar details (for a full list of annotated features see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and,'' ``or,'' ``but,'' etc.) across all features and scenes as the ``vocabulary'' for the topic model.  We then concatenated the sets of words across all features contained in overlapping, 50-scene sliding windows, and treated each 50-scene sequence as a single ``document'' for the purpose of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 27 unique topics (with non-zero weights) were sufficient to describe the time-varying content of the video (see \textit{Methods}; Figs.~\ref{fig:schematic}, \topics).  Note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06}, in that we sought to characterize how the thematic content of the episode evolved over time.  However, whereas Dynamic Topic Models are designed to characterize how the properties of \textit{collections} of documents change over time, our sliding window approach allows us to examine the topic dynamics within a single document (or video).  Specifically, our approach yielded (via the topic proportions matrix) a single \textit{topic vector} for each timepoint of the episode (we set timepoints to match the acquisition times of the 1976 fMRI volumes collected as participants viewed the episode).

@@

The topics we found were heavily character-focused (e.g., the top-weighted word in each topic was nearly always a character) and could be roughly divided into themes that were primarily Sherlock Holmes-focused (Sherlock is the titular character); primarily John Watson-focused (John is Sherlock's close confidant and assistant); or that involved Sherlock and John interacting (Fig.~\topics).  Several of the topics were highly similar, which we hypothesized might allow us to distinguish between subtle narrative differences (if the distinctions between those overlapping topics were meaningful; also see Fig.~\featureimportance).  The topic vectors for each timepoint were \textit{sparse}, in that only a small number (usually one or two) of topics tended to be ``active'' in any given timepoint (Fig.~\ref{fig:model}A).  Further, the dynamics of the topic activations appeared to exhibit \textit{persistance} (i.e., given that a topic was active in one timepoint, it was likely to be active in the following timepoint) along with \textit{occasional rapid changes} (i.e., occasionally topics would appear to spring into or out of existence).  These two properties of the topic dynamics may be seen in the block diagonal structure of the timepoint-by-timepoint correlation matrix (Fig.~\ref{fig:model}B) and reflect the gradual drift and sudden shifts fundamental to the contextual dynamics of real-world experiences.  Given this observation, we adapted an approach devised by \cite{BaldEtal17}, and used a Hidden Markov Model (HMM) to identify the \textit{event boundaries} where the topic activations changed rapidly (i.e., at the boundaries of the blocks in the correlation matrix; event boundaries identified by the HMM are outlined in yellow).  Part of our model fitting procedure required selecting an appropriate number of ``events'' to segment the timeseries into.  We used an optimization procedure to identify the number of events that maximized within-event stability while also minimizing across-event correlations (see \textit{Methods} for additional details).  To create a stable ``summary'' of the video, we computed the average topic vector within each event (Fig.~\ref{fig:model}C).

@@

Given that the time-varying content of the video could be segmented cleanly into discrete events, we wondered whether participants' recalls of the video also displayed a similar structure.  We applied the same topic model (already trained on the video annotations) to each participant's recalls.  Analogous to how we analyzed the time-varying content of the video, to obtain similar estimates for participants' recalls, we treated each (overlapping) 10 sentence ``window'' of their transcript as a ``document'' and then computed the most probable mix of topics reflected in each timepoint's sentences.  This yielded, for each participant, a number-of-sentences by number-of-topics topic proportions matrix that characterized how the topics identified in the original video were reflected in the participant's recalls.  Note that an important feature of our approach is that it allows us to compare participant's recalls to events from the original video, despite that different participants may have used different language to describe the same event, and that those descriptions may not match the original annotations.  This is a substantial benefit of projecting the video and recalls into a shared ``topic'' space.  An example topic proportions matrix from one participant's recalls is shown in Figure~\ref{fig:model}D.

@@

Two clear patterns emerged from this set of analyses.  First, although every individual participant's recalls could be segmented into discrete events (i.e., every individual participant's recall correlation matrix exhibited clear block diagonal structure; Fig.~\corrmats), each participant appeared to have a unique \textit{recall resolution}, reflected in the sizes of those blocks.  For example, some participants' recall topic proportions segmented into just a few events (e.g., Participants P1, P4, and P15), while others' recalls segmented into many shorter duration events (e.g., Participants P12, P13, and P17).  This suggests that different participants may be recalling the video with different levels of detail-- e.g., some might touch on just the major plot points, whereas others might attempt to recall every minor scene or action.  The second clear pattern present in every individual participant's recall correlation matrix is that, unlike in the video correlation matrix, there are substantial off-diagonal correlations in participant's recalls.  Whereas each event in the original video was (largely) separable from the others (Fig.~\ref{fig:model}B), in transforming those separable events into memory participants appear to be integrating \textit{across} different events, blending elements of previously recalled and not-yet-recalled events into each newly recalled event~\citep[Figs.~\ref{fig:model}D, \corrmats; also see][]{MannEtal11, HowaEtal12}.

The above results indicate that both the structure of the original video and participants' recalls of the video exhibit event boundaries that can be identified automatically by characterizing the dynamic content using a shared topic model and segmenting the content into events using HMMs.  Next, we asked whether some correspondence might be made between the specific content of the events the participants experienced in the video, and the events they later recalled.  One approach to linking the experienced (video) and recalled events is to label each recalled event as matching the video event with the most similar (i.e., most highly correlated) topic vector (Figs.~\ref{fig:model}G, \matchmats).  This yields a sequence of ``presented'' events from the original video, and a sequence of (potentially differently ordered) ``recalled'' events for each participant.  Analogous to classic list-learning studies, we can then examine participants' recall sequences by asking which events they tended to recall first~\citep[probability of first recall; Fig.~\ref{fig:list-learning}A;][]{WelcBurn24, PostPhil65, AtkiShif68}; how participants most often transition between recalls of the events as a function of the temporal distance between them~\citep[lag-conditional response probability; Fig.~\ref{fig:list-learning}B;][]{Kaha96}; and which events they were likely to remember overall~\citep[serial position recall analyses; Fig.~\ref{fig:list-learning}C;][]{Murd62a}. Interestingly, for two of these analyses (probability of first recall and lag-conditional response probability curves) we observe patterns comparable to classic effects from the list-learning literature: namely, a higher probability of initiating recall with the first event in the sequence (Fig.~\ref{fig:list-learning}A) and a higher probability of transitioning to neighboring events with an asymmetric forward bias (Fig.~\ref{fig:list-learning}C). In contrast, we do not observe a pattern comparable to the serial position effect (Fig.~\ref{fig:list-learning}C), but rather we see higher memory for specific events distributed somewhat evenly throughout the video.

We can also apply two analyses native to list-learning studies that describe how participants group items in their recall sequences: temporal clustering and semantic clustering \citep[][see \textit{Methods} for details]{PolyEtal09}.  Temporal clustering refers to the extent to which participants group their recall responses according to encoding position.  Overall, we found that sequentially viewed video events were clustered heavily in participants' recall event sequences (mean: 0.808, SEM: 0.022) and that participants with higher temporal clustering scores also tended to recall more events (Pearson's $r(15) = 0.62, p = 0.007$).  Semantic clustering measures the extent to which participants cluster their recall responses according to semantic similarity.  We found that participants tended to recall semantically similar video events together (mean: 0.813, SEM: 0.022) and participants who exhibited stronger semantic clustering scores overall remembered more video events (Pearson's $r(15) = 0.55, p = 0.02$).

@@

Statistical models of memory studies often treat memory recalls as binary (e.g. the item was recalled or not) and independent events.  However, our framework produces a content-based model of individual stimulus and recall events, allowing for direct quantitative comparison between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel metrics for quantifying naturalistic memory representations:~\textit{precision} and \textit{distinctiveness}.  We define precision as the average correlation between the topic proportions of each recall event and the maximally correlated video event (Fig.~\ref{fig:precision-distinctiveness}).  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that across participants, a higher precision score is correlated to both hand annotated memory performance (Pearson's $r(15) = 0.6, p = 0.011$) and the number of recall events estimated by our model (Pearson's $r(15) = 0.64, p = 0.005$). A second novel metric we introduce here is distinctiveness, or how unique the recall description was to each video event. We define distinctiveness as 1 minus the average of all non-matching recall events from the video-recall correlation matrix. We hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  Similarly to precision, we find that the more distinct participants recalls are (on average), the more they remembered (hand-annotated memory: Pearson's $r(15) = 0.83, p < 0.001$; model-derived memory: Pearson's $r(15) = 0.71, p = 0.001$).  In summary, using two novel metrics afforded by our approach, we find that participants whose recalls are both more precise and distinct remember more content.

@@

Thus, analyses such as those in Figure \ref{fig:list-learning} that address only the temporal dynamics of free recall paint an incomplete picture of memory for naturalistic episodes.  While useful for studying presentation order-dependent recall dynamics, they neglect to consider the stimuli's content (or, for example, that content's potential interrelatedness).  However, sensitivity to stimulus and recall content introduces a new challenge: distinguishing between levels of recall quality for a stimulus (i.e., an event) that is considered to have been ``remembered."  When modeling memory experiments, often times events (or items) and their later memories are treated as binary and independent events (e.g., a given list item was simply either remembered or not remembered).  Various models of memory (e.g., \citealp{Yone02}) attempt to improve upon this by including confidence ratings, rendering the binary judgement instead categorical.  Our novel framework allows one to assess memory performance in a more continuous way (\textit{precision}), as well as analyze the correlational structure of each encoding event to each memory event (\textit{distinctiveness}). Further and importantly, these two novel metrics we introduce here arise from comparisons of the actual content of the experience/memories, which is not typically modeled.  Leveraging this, we find that the successful memory performance is related to 1) the precision with which the participant recounts each event and 2) the distinctiveness of each recall event (relative to the other recalled events). The first finding suggests that the information retained for \textit{any individual event} may predict the overall amount of information retained by the participant.  The second finding suggests that the ability to distinguish between temporally or semantically similar content is also related to the quantity of content recovered. Intriguingly, prior studies show that pattern separation, or the ability to discriminate between similar experiences, is impaired in many cognitive disorders as well as natural aging \citep{StarEtal10, YassEtal11c, YassStar11b}.  Future work might explore whether and how these metrics compare between cognitively impoverished groups and healthy controls.

>>>>

%\title{How is experience transformed into memory?}
%\title{A content-based framework for capturing stimulus and memory dynamics reveals event-like structure in naturalistic episodic recall}
%\title{A content-based model characterizing/describing/representing how experiences and memories unfold over time reveals event-like structure in naturalistic episodic recall}
%\title{A content-based framework describing the temporal dynamics of experiences and memories reveals event-like structure in naturalistic episodic recall}
%\title{A content-based model of how experiences and memories unfold over time reveals event-like structure in naturalistic episodic recall}
%\title{Recalling naturalistic experiences exhibits event-like structure}
%\title{A novel framework for linking dynamic experiences to memories reveals event-like structure in naturalistic episodic recall}
\title{A content-sensitive framework for modeling episodic memory reveals event-like structure in naturalistic experience, recall, and neural processing}

@@

\begin{abstract}
Our life experiences unfold over time in highly complex manner, with the evolving presence of numerous intricate features describing our journey between each event we encounter. Here, we propose a framework for mapping dynamic naturalistic episodes onto geometric spaces as \textit{experience trajectories} that capture the temporal dynamics of real-world content. Within this geometric framework, one may compare the shape of the trajectory formed by an experience to that defined by one's later recollection to characterize our memories' recovery and distortion of the external world. Here, we apply this approach to a naturalistic memory experiment in which participants viewed and verbally recounted a video, and find that the video and subsequent recalls share both an experience-specific shape and a discernible event-like structure.  Despite this apparent similarity, we find that the level of \textit{precision} with which individuals recounted various events and the \textit{distinctiveness} of recall for those events were varied and predictive of overall memory performance. Finally, we identify a network of brain structures that is sensitive to the ``shapes" of our ongoing experiences, and an overlapping network sensitive to how we will later remember them.  Our framework and findings underscore the rich event-like structure of the external world and our memories, and offer novel, content-sensitive measures for assessing episodic memory
%Our life experiences unfold over time and the temporal dynamics of their contents form unique \textit{experience trajectories}.  Within this geometric framework, one can compare the shape of the trajectory formed by an experience to that defined by our later remembering of that experience.  We propose a framework for mapping naturalistic experiences onto geometric spaces that characterize how they unfold over time.  We apply this approach to a naturalistic memory experiment which had participants view and recount a video.  We find that the video and subsequent recall share an event-like structure, and that the ``shapes'' of the trajectories formed by participants' recountings were all highly similar to that of the original video. Further, participants' that recounted the events with a high level of precision and in a distinctive way had better subsequent memory performance. Lastly, we identified a network of brain structures that are sensitive to the shapes of our ongoing experiences, and an overlapping network that is sensitive to how we will later remember those experiences. These results highlight the rich event-like structure of memories for our life experiences, and introduce a novel framework for mapping between dynamic life experiences and their mnemonic counterparts.
\end{abstract}

@@

What does it mean to \textit{remember} something? In traditional episodic memory experiments \citep[e.g., list-learning or trial-based experiments;][]{Murd62a, Kaha96}, remembering is often cast as a discrete and binary operation: each studied item may be separated from all others, and labeled as having been recalled or forgotten. More nuanced studies might incorporate self-reported confidence measures as a proxy for memory strength, or ask participants to discriminate between ``recollecting'' the (contextual) details of an experience or having a general feeling of ``familiarity'' \citep{Yone02}. Using well-controlled, trial-based experimental designs, the field has amassed a wealth of valuable information regarding human episodic memory.  However, there are fundamental properties of the external world and our memories that trial-based experiments are not well suited to capture~\citep[for review also see][]{KoriGold94, HukEtal18}.  First, our experiences and memories are continuous, rather than discrete---removing a (naturalistic) event from the context in which it occurs can substantially change its meaning.  Second, the specific language used to describe an experience has little bearing on whether the experience should be considered to have been ``remembered.''  Asking whether the rememberer has precisely reproduced a specific set of words to describe a given experience is nearly orthogonal to whether they were actually able to remember it.  In classic (e.g., list-learning) memory studies, by contrast, the number or proportion of precise recalls is often a primary metric for assessing the quality of participants' memories.  Third, one might remember the \textit{essence} (or a general summary) of an experience but forget (or neglect to recount) particular details.  Capturing the essence of what happened is typically the main ``point'' of recounting a memory to a listener, while the addition of highly specific details may add comparatively little to successful conveyance of an experience.

How might one go about formally characterizing the ``essence" of an experience, or whether it has been recovered by the rememberer?  Any given moment of an experience derives meaning from surrounding moments, as well as from longer-range temporal associations~\citep{LernEtal11, Mann19}.  Therefore, the timecourse describing how an event unfolds is fundamental to its overall meaning.  Further, this hierarchy formed by our subjective experiences at different timescales defines a \textit{context} for each new moment~\citep[e.g., ][]{HowaKaha02a, HowaEtal14}, and plays an important role in how we interpret that moment and remember it later~\citep[for review see][]{MannEtal15}.  Our memory systems can leverage these associations to form predictions that help guide our behaviors~\citep{RangRitc12}.  For example, as we navigate the world, the features of our subjective experiences tend to change gradually (e.g., the room or situation we are in at any given moment is strongly temporally autocorrelated), allowing us to form stable estimates of our current situation and behave accordingly~\citep{ZackEtal07, ZwaaRadv98}.

@@

Here we sought to examine how the temporal dynamics of a ``naturalistic'' experience were later reflected in participants' memories.  We analyzed an open dataset that comprised behavioral and functional Magnetic Resonance Imaging (fMRI) data collected as participants viewed and then verbally recounted an episode of the BBC television series \textit{Sherlock}~\citep{ChenEtal17}.  We developed a computational framework for characterizing the temporal dynamics of the moment-by-moment content of the episode and of participants' verbal recalls.  Specifically, we use topic modeling~\citep{BleiEtal03} to characterize the thematic conceptual (semantic) content present in each moment of the episode and recalls, and Hidden Markov Models~\citep{Rabi89, BaldEtal17} to discretize this evolving semantic content into events.  In this way, we cast naturalistic experiences (and recalls of those experiences) as \textit{trajectories} that describe how the experiences evolve over time. Under this framework, successful remembering entails verbally ``traversing'' the content trajectory of the episode, thereby reproducing the shape (or essence) of the original experience.  Comparing the shapes of the topic trajectories of the episode and of participants' retellings of the episode then reveals which aspects of the episode were preserved (or lost) in the translation into memory.  We further examine whether 1) the \textit{precision} with which a participant recounts each event and 2) the \textit{distinctiveness} each recall event is (relative to the other recalled events) relates to their overall memory performance.  Last, we identify networks of  brain structures whose responses (as participants watched the episode) reflected the temporal dynamics of the episode, and how participants would later recount the episode.

\section*{Results}
To characterize the shape of the \textit{Sherlock} episode and participants' subsequent recountings of its unfolding, we used a topic model~\citep{BleiEtal03} to discover the latent themes in the episode's dynamic content.  Topic models take as inputs a vocabulary of words to consider and a collection of text documents, and return two output matrices.  The first of these is a \textit{topics matrix} whose rows are topics (latent themes) and whose columns correspond to words in the vocabulary. The entries of the topics matrix define how each word in the vocabulary is weighted by each discovered topic.  For example, a detective-themed topic might weight heavily on words like ``crime,'' and ``search.''  The second output is a \textit{topic proportions matrix}, with one row per document and one column per topic.  The topic proportions matrix describes what mixture of discovered topics is reflected in each document.

\cite{ChenEtal17} collected hand-annotated information about each of 1000 (manually identified) scenes spanning the roughly 50 minute video used in their experiment.  This information included: a brief narrative description of what was happening; whether the scene took place indoors or outdoors; the names of any characters on the screen; the names of any characters who were in focus in the camera shot; the names of characters who were speaking; the location where the scene took place; the camera angle (close up, medium, long, etc.); whether or not background music was present; and other similar details (for a full list of annotated features see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and,'' ``or,'' ``but,'' etc.) across all features and scenes as the ``vocabulary'' for the topic model.  We then concatenated the sets of words across all features contained in overlapping, 50-scene sliding windows, and treated each 50-scene sequence as a single ``document'' for the purpose of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 27 unique topics (with non-zero weights) were sufficient to describe the time-varying content of the video (see \textit{Methods}; Figs.~\ref{fig:schematic}, \topics).  Note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06} in that we sought to characterize how the thematic content of the episode evolved over time.  However, whereas Dynamic Topic Models are designed to characterize how the properties of \textit{collections} of documents change over time, our sliding window approach allows us to examine the topic dynamics within a single document (or video).  Specifically, our approach yielded (via the topic proportions matrix) a single \textit{topic vector} for each timepoint of the episode (we set timepoints to match the acquisition times of the 1976 fMRI volumes collected as participants viewed the episode).

@@

The topics we found were heavily character-focused (e.g., the top-weighted word in each topic was nearly always a character) and could be roughly divided into themes that were primarily Sherlock Holmes-focused (Sherlock is the titular character), primarily John Watson-focused (John is Sherlock's close confidant and assistant), or focused on Sherlock and John interacting (Fig.~\topics).  Several of the topics were highly similar, which we hypothesized might allow us to distinguish between subtle narrative differences (if the distinctions between those overlapping topics were meaningful; also see Fig.~\featureimportance).  The topic vectors for each timepoint were \textit{sparse}, in that only a small number (usually one or two) of topics tended to be ``active'' in any given timepoint (Fig.~\ref{fig:model}A).  Further, the dynamics of the topic activations appeared to exhibit \textit{persistance} (i.e., given that a topic was active in one timepoint, it was likely to be active in the following timepoint) along with \textit{occasional rapid changes} (i.e., occasionally topics would appear to spring into or out of existence).  These two properties of the topic dynamics may be seen in the block diagonal structure of the timepoint-by-timepoint correlation matrix (Fig.~\ref{fig:model}B) and reflect the gradual drift and sudden shifts fundamental to the contextual dynamics of real-world experiences.  Given this observation, we adapted an approach devised by \cite{BaldEtal17}, and used a Hidden Markov Model (HMM) to identify the \textit{event boundaries} where the topic activations changed rapidly (i.e., at the boundaries of the blocks in the correlation matrix; event boundaries identified by the HMM are outlined in yellow).  Part of our model fitting procedure required selecting an appropriate number of ``events'' to segment the timeseries into.  We used an optimization procedure to identify the number of events that maximized within-event stability while also minimizing across-event correlations (see \textit{Methods} for additional details).  To create a stable ``summary'' of the video, we computed the average topic vector within each event (Fig.~\ref{fig:model}C).

@@

Given that the time-varying content of the video could be segmented cleanly into discrete events, we wondered whether participants' recalls of the video also displayed a similar structure.  We applied the same topic model (already trained on the video annotations) to each participant's recalls.  Analogous to how we analyzed the time-varying content of the video, to obtain similar estimates for participants' recalls, we treated each (overlapping) 10-sentence ``window'' of their transcript as a ``document'' and then computed the most probable mix of topics reflected in each timepoint's sentences.  This yielded, for each participant, a number-of-sentences by number-of-topics topic proportions matrix that characterized how the topics identified in the original video were reflected in the participant's recalls.  Note that an important feature of our approach is that it allows us to compare participant's recalls to events from the original video, despite that different participants may have used different language to describe the same event, and that those descriptions may not match the original annotations.  This is a substantial benefit of projecting the video and recalls into a shared ``topic'' space.  An example topic proportions matrix from one participant's recalls is shown in Figure~\ref{fig:model}D.

@@

Two clear patterns emerged from this set of analyses.  First, although every individual participant's recalls could be segmented into discrete events (i.e., every individual participant's recall correlation matrix exhibited clear block diagonal structure; Fig.~\corrmats), each participant appeared to have a unique \textit{recall resolution}, reflected in the sizes of those blocks.  For example, some participants' recall topic proportions segmented into just a few events (e.g., Participants P1, P4, and P15), while others' recalls segmented into many shorter duration events (e.g., Participants P12, P13, and P17).  This suggests that different participants may be recalling the video with different levels of detail-- e.g., some might touch on just the major plot points, whereas others might attempt to recall every minor scene or action.  The second clear pattern present in every individual participant's recall correlation matrix is that, unlike in the video correlation matrix, there are substantial off-diagonal correlations.  Whereas each event in the original video was (largely) separable from the others (Fig.~\ref{fig:model}B), in transforming those separable events into memory, participants appear to be integrating across mutliple events, blending elements of previously recalled and not-yet-recalled events into each newly recalled event~\citep[Figs.~\ref{fig:model}D, \corrmats; also see][]{MannEtal11, HowaEtal12}.

The above results indicate that both the structure of the original video and participants' recalls of the video exhibit event boundaries that can be identified automatically by characterizing the dynamic content using a shared topic model and segmenting the content into events using HMMs.  Next, we asked whether some correspondence might be made between the specific content of the events the participants experienced in the video, and the events they later recalled.  One approach to linking the experienced (video) and recalled events is to label each recalled event as matching the video event with the most similar (i.e., most highly correlated) topic vector (Figs.~\ref{fig:model}G, \matchmats).  This yields a sequence of ``presented'' events from the original video, and a (potentially differently ordered) sequence of ``recalled'' events for each participant.  Analogous to classic list-learning studies, we can then examine participants' recall sequences by asking which events they tended to recall first~\citep[probability of first recall; Fig.~\ref{fig:list-learning}A;][]{WelcBurn24, PostPhil65, AtkiShif68}; how participants most often transition between recalls of the events as a function of the temporal distance between them~\citep[lag-conditional response probability; Fig.~\ref{fig:list-learning}B;][]{Kaha96}; and which events they were likely to remember overall~\citep[serial position recall analyses; Fig.~\ref{fig:list-learning}C;][]{Murd62a}. Interestingly, for two of these analyses (probability of first recall and lag-conditional response probability curves) we observe patterns comparable to classic effects from the list-learning literature: namely, a higher probability of initiating recall with the first event in the sequence (Fig.~\ref{fig:list-learning}A) and a higher probability of transitioning to neighboring events with an asymmetric forward bias (Fig.~\ref{fig:list-learning}C). In contrast, we do not observe a pattern comparable to the serial position effect (Fig.~\ref{fig:list-learning}C), but rather we see higher memory for specific events distributed somewhat evenly throughout the video.

We can also apply two list-learning-native analyses that describe how participants group items in their recall sequences: temporal clustering and semantic clustering \citep[][see \textit{Methods} for details]{PolyEtal09}.  Temporal clustering refers to the extent to which participants group their recall responses according to encoding position.  Overall, we found that sequentially viewed video events were clustered heavily in participants' recall event sequences (mean: 0.808, SEM: 0.022) and that participants with higher temporal clustering scores also tended to recall more events (Pearson's $r(15) = 0.62, p = 0.007$).  Semantic clustering measures the extent to which participants cluster their recall responses according to semantic similarity.  We found that participants tended to recall semantically similar video events together (mean: 0.813, SEM: 0.022) and participants who exhibited stronger semantic clustering remembered more video events overall (Pearson's $r(15) = 0.55, p = 0.02$).

@@

Statistical models of memory studies often treat memory recalls as binary (e.g. the item was recalled or not) and independent events.  However, our framework produces a content-based model of individual stimulus and recall events, allowing for direct quantitative comparison between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel metrics for quantifying naturalistic memory representations:~\textit{precision} and \textit{distinctiveness}.  We define precision as the average correlation between the topic proportions of each recall event and the maximally correlated video event (Fig.~\ref{fig:precision-distinctiveness}).  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that, across participants, a higher precision score is correlated to both hand-annotated memory performance (Pearson's $r(15) = 0.6, p = 0.011$) and the number of recall events estimated by our model (Pearson's $r(15) = 0.64, p = 0.005$). A second novel metric we introduce here is distinctiveness, or how unique the recall description was to each video event. We define distinctiveness as 1 minus the average of all non-matching recall events from the video-recall correlation matrix. We hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  Similarly to precision, we find that the more distinct participants recalls are (on average), the more they remembered (hand-annotated memory: Pearson's $r(15) = 0.83, p < 0.001$; model-derived memory: Pearson's $r(15) = 0.71, p = 0.001$).  In summary, using two novel metrics afforded by our approach, we find that participants whose recalls are both more precise and distinct remember more content.

@@

Thus, analyses such as those in Figure \ref{fig:list-learning} that address only the temporal dynamics of free recall paint an incomplete picture of memory for naturalistic episodes.  While useful for studying presentation order-dependent recall dynamics, they neglect to consider the stimuli's content (or, for example, that content's potential interrelatedness).  However, sensitivity to stimulus and recall content introduces a new challenge: distinguishing between levels of recall quality for a stimulus (i.e., an event) that is considered to have been ``remembered."  When modeling memory experiments, often times events (or items) and their later memories are treated as binary and independent events (e.g., a given list item was simply either remembered or not remembered).  Various models of memory (e.g., \citealp{Yone02}) attempt to improve upon this by including confidence ratings, rendering this binary judgement instead categorical.  Our novel framework allows one to assess memory performance in a more continuous way (\textit{precision}), as well as analyze the correlational structure of each encoding event to each memory event (\textit{distinctiveness}). Further and importantly, these two novel metrics we introduce here arise from comparisons of the actual content of the experience/memories, which is not typically modeled.  Leveraging this, we find that the successful memory performance is related to 1) the precision with which the participant recounts each event and 2) the distinctiveness of each recall event (relative to the other recalled events). The first finding suggests that the information retained for \textit{any individual event} may predict the overall amount of information retained by the participant.  The second finding suggests that the ability to distinguish between temporally or semantically similar content is also related to the quantity of content recovered. Intriguingly, prior studies show that pattern separation, or the ability to discriminate between similar experiences, is impaired in many cognitive disorders as well as natural aging \citep{StarEtal10, YassEtal11c, YassStar11b}.  Future work might explore whether and how these metrics compare between cognitively impoverished groups and healthy controls.
